---
title: 'Etude de cas *Titanic*'
subtitle: "Résultats statistiques bruts"
author:  "Bruno Fischer Colonimos"
date: "`r format(Sys.Date(), '%d %B %Y')`"
# header-includes: \usepackage{graphicx} # for unmodded template
header-includes: \usepackage{float} # for fig.pos = 'H'
fontsize: 12pt
urlcolor: "blue"   ## external
linkcolor: "red" ## internal
documentclass: article
classoption: a4paper
geometry: "left=1.5cm,right=1.5cm,top=1.5cm,bottom=2cm,footskip=1cm"
footer: \thetitle \ \textemdash \ \thepage ## verify this (theTitle == thetitle)
output:
  pdf_document:
    # template: test_latex_temp
    template: latex_temp2.latex
    highlight: monochrome
    number_sections: yes
    toc: yes
    toc_depth: 4
    # keep_tex: true
    fig_caption: yes
    keep_tex: yes

  # html_document:
  #   number_sections: yes
  #   toc: yes
  #   toc_depth: 4
  #   fig_caption: yes
---



<!-- Some Inline HTML (CSS) , formatting tables -->

<style>

table {
    width: 100%;
    border-top-style: solid;
    border-top-width: medium;
    border-top-color: initial;
    border-bottom-style: solid;
    border-bottom-width: medium;
    border-bottom-color: initial;
}

</style>

<!-- End Inline HTML -->




```{r setup0, include=FALSE}

# Preliminary code
# ================
## Libraries and auxiliary code (install before running)

### code config parameters

# adjust Gggplot2 theme and color palette
bwtheme <- TRUE
specialpalette <- FALSE
showarning <- FALSE
datmis <- "Ukn"


### Required packages (install before running)
# libs_prep, echo =FALSE, warning=FALSE, message=FALSE, results='hide'}

library(knitr)
library(pander)
# library(caret)
library(ggplot2)
library(dplyr)
library(grid)
library(gridExtra)

# Source preparatory code "preparation.R"
# ========================

## code config parameters
# adjust (?) Gggplot2 theme and color palette
bwtheme <- TRUE         # use bw theme
specialpalette <- FALSE # use the special palette for color-blinds
showarning <- FALSE     # show the warnings
datmis <- "Ukn"         # Value to replace the NA's with 
# Execution
echo_code <- TRUE       # Echo the code chunks
outresults <- TRUE      # Output the results (printing, etc)
saveresults <- TRUE     # save the results in Rda file for reuse in report


source("preparation.R")

# Set aliases

knitr::set_alias(w = "fig.width", a = "fig.asp")

```

```{r setup, echo = FALSE}
# setup chunk : default options
knitr::opts_chunk$set(echo = FALSE, fig.width = w.13, fig.asp = a.11 , fig.pos = 'H', fig.align = "center", fig.show = "hold", warning = showarning)
```

```{r morecode}
displaygraph <- function(one.graph,
                         windevice = NULL,
                         width = 6,
                         height = 5 ) {
        if (is.null(windevice)) {
                grid.newpage()
                grid.draw(one.graph)
        } else {
                windows(width = width, height = height)
                grid.newpage()
                grid.draw(one.graph)
                Sys.sleep(defsleep)
                dev.off()
        }
}

# for display several ggplots or lists of ggplots as one arrangeGrob call
# 
# moklist <- function(lisobj) {
#         test <- function(obj) {"gg" %in% class(obj)}
#         
#         unlist(
#                 lapply(lisobj, 
#                        FUN = function(x){
#                                if (test(x)){
#                                        list(x)
#                                }else{x}
#                        }
#                 ),
#                 recursive = FALSE
#         )
# }


moklist <- function(...) { # arguments may be gg_graphs or lists of graphs
        test <- function(obj) {"gg" %in% class(obj)}
        lisobj <- list(...)
        
        unlist(
                lapply(lisobj, 
                       FUN = function(x){
                               if (test(x)){
                                       list(x)
                               }else{x}
                       }
                ),
                recursive = FALSE
        )
}



# Functions for showing results in Markdown text

# show proportion as %
as.percent <- function(prop, digits = 0, suffix = "%", sep = " ") 
        paste(100 * round(prop, digits = digits + 2), 
              suffix, sep = sep)


# local assignments for  displaying values in the text

loc <- list(
        var = list(),
        let = function(...) {loc$var <<- list(...)},
        flush = function() loc$var <<-list()
        )

## usage
# loc$var # initial state of values
# loc$let(a = 3, b = 5) # set values
# loc$var # get list key-value
# loc$var$a # get value
# loc$flush() # erase values
# loc$var
# loc$var$a


```





```{r data}
# get the analyzed data
savedir <- "data"
savefname <- "analysis_results.rda"
load(file=file.path(savedir, savefname))
```

--------------------------------------------------------------------------------------------

Les Passagers
=============


Sexe
----

```{r anx_sex}
pander(tb_sex)
displaygraph(p_sex)

```

## Age, Statut lié à l'âge (Age, Agestatus)

```{r anx_age, fig.width=w.11, fig.asp=a.12}
pander(tb_age)

local({
        lgrob = moklist(p_age, p_age_sex, p_agestat_by )
        gr <- arrangeGrob(
                grobs = lgrob,
                ncol = 3)
        displaygraph(gr)
})

```

## Groupe familial

Remarque : la variable Famly a été dréée par l'addition des variables Sibsp et Parch

```{r anx_famly, fig.show = "hold", fig.width=w.13, fig.asp=a.11}
p_famly$famly

pander(rbind(tb_class_by$gsize, tb_class_by$gsize_p) , caption = "")
```

## Classes de tarif (Pclass)

```{r anx_pclass}
pander(tb_pclass)
p_pclass + labs( title = "")

```

```{r pclass_by, fig.width = w.11, fig.asp = a.11}

pander(tb_class_by)

local({
        gr <- arrangeGrob(
                grobs = p_age_by,
                ncol = 3)
        displaygraph(gr)
})
```


### Classes de tarif et placement dans le navire

```{r fig.width=w.11, fig.asp=a.13}
local({
        gr <- arrangeGrob(
                grobs = p_deck,
                ncol = 3)
        displaygraph(gr)
})

```


## Port d'embarquement


```{r fig.width=w.11, fig.asp= a.12}
pander(tb_embark)

local({
        gr <- arrangeGrob(
                grobs = p_embark,
                ncol = 3)
        displaygraph(gr)
})

```


## Prix du voyage (Fare)

```{r fig.width=w.11, fig.asp= a.12}
pander(tb_fare)

local({
        gr <- arrangeGrob(
                grobs = p_fare,
                ncol = 3)
        displaygraph(gr)
})

```



La survie
==========


## Survie

```{r survie, fig.width= w.13, fig.asp= a.11}
pander(tb_survived)

p_survived

# local({
#         gr <- arrangeGrob(
#                 p_survived,
#                 ncol = 3)
#         displaygraph(gr)
# })

```

## Survie et caractéristiques démographiques

```{r fig.width=w.11, fig.asp= a.11}
pander(tb_survived_by, digits=2,
                       caption = "Survival by Sex , class")
local({
        gr <- arrangeGrob(
                grobs = p_survived_by,
                ncol = 3)
        displaygraph(gr)
})
```



```{r fig.width=w.11, fig.asp= a.13}
# pander(tb_survived_by, digits=2,
#                        caption = "Survival by Sex , class")
local({
        gr <- arrangeGrob(
                grobs = p_age_survived,
                ncol = 3)
        displaygraph(gr)
})
```



```{r fig.width=w.11, fig.asp= a.34}
pander(survived_age_sex_class_f_p, digits=2, 
                        caption = "Survival by  Class, Sex , Age ")

local({
        gr <- arrangeGrob(
                grobs = pl_survived_by,
                ncol = 3)
        displaygraph(gr)
})

```
                        




