---
title: 'Titanic - Kaggle - Some exploration'
subtitle: "Explore for classroom use"
author:  "Bruno Fischer Colonimos"
date: "`r format(Sys.Date(), '%d %B %Y')`"
# header-includes: \usepackage{graphicx} # for unmodded template
fontsize: 12pt
urlcolor: "blue"   ## external
linkcolor: "red" ## internal
documentclass: article
classoption: a4paper
geometry: "left=1.5cm,right=1.5cm,top=1.5cm,bottom=2cm,footskip=1cm"
footer: \thetitle \ \textemdash \ \thepage 
output:
  # bookdown::pdf_document2:
  pdf_document:
    # template: test_latex_temp
    template: latex_temp2.latex
    highlight: monochrome
    number_sections: yes
    toc: yes
    toc_depth: 4
    # keep_tex: true
    fig_caption: yes
    
  # html_document:
  #   number_sections: yes
  #   toc: yes
  #   toc_depth: 4
---

----------------


<!-- Some Inline HTML (CSS) , formatting tables -->

<style>

table {
    width: 100%;
    border-top-style: solid;
    border-top-width: medium;
    border-top-color: initial;
    border-bottom-style: solid;
    border-bottom-width: medium;
    border-bottom-color: initial;
}

</style>

<!-- End Inline HTML -->



Preliminary code
================


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.keep="all" ,fig.width=4, fig.asp= 0.75 , fig.align="center", fig.keep ="all", fig.show="hold", out.width="0.48\\textwidth")

# , fig.align="center"
# not for HTML : out.width="0.48\\textwidth"
```


## Required packages (install before running)

"caret", "ggplot2", pander       



## Libraries and auxiliary code (install before running)

(Not fully echoed here)

```{r}
# adjust Gggplot2 theme and color palette
bwtheme <- TRUE
specialpalette <- TRUE # controls the colour of the plots
```


```{r libs, echo =FALSE, warning=FALSE, message=FALSE, results='hide'}
library(knitr)
library(pander)
library(caret)
library(ggplot2)
# library(party) #trees
# library(ISLR) # example
# 
# library(ElemStatLearn)
if (bwtheme) {
        theme_set(theme_bw())
}

if (specialpalette) {
        # palette color-blind-friendly: The palette with black (Cookbook for R,
        # http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/)
        cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73",
                        "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
        
        # modified BFC
        cbfPalette <- cbbPalette
        cbfPalette[3] <- cbbPalette[7]
        cbfPalette[6] <- cbbPalette[3]
        cbfPalette[8] <- cbbPalette[7]
        
        # * `set_default_scale` has been removed. If you want to change the default
        # scale for an aesthetic, just create a function called
        # `scale_aesthetic_continuous` or `scale_aesthetic_discrete` that returns the
        # scale that you want. 
        
        scale_colour_discrete <- function(...) {
                scale_colour_manual(values = cbbPalette, ...)
        }
        
        scale_fill_discrete <- function(...) {
                scale_fill_manual(values = cbbPalette, ...)
        }
        
        scale_colour_continuous <- function(...) {
                ggplot2::scale_colour_brewer("Greys", ...)
        }
        
}

```

Data
=====

```{r dataget}

datadir <-  "data"
fname <-  "train.csv"

tdf <- read.csv(file.path(datadir, fname))

# get the chunk label
chlabel <- knitr::opts_current$get("label")# does'nt work
# verify
paste0( "chlabel =", chlabel)
# alternatives
# chlabel <- knitr::opts_current$get()$label
# paste0( "chlabel =", chlabel)
# 
# copts <- knitr::opts_current$get()
# copts

# tabcap <- function(caption, chunklabel = chlabel) {
#         paste0("(#tab:", chunklabel, ")", caption)
# }

tabcap <- function(caption, chunklabel=knitr::opts_current$get("label")) {
        paste0("(#tab:", chunklabel, ")", caption)
}

# pander(head(tdf), caption = "\\label{tab:dataget}A `Pander` table")
# knitr::kable(head(tdf), caption = "A `knitr` table")
# pander(head(tdf), caption = "(#tab:dataget) A `Pander` table")
pander(head(tdf), caption = tabcap("A `Pander` table"))     
```

* Quick dataframe summary
```{r datasum}
# pander(summary(tdf), caption="( )Data Head")
# knitr::kable(summary(tdf), caption="Data Head")
# pander(summary(tdf), caption="(#tab:datasum)Data summaries")

pander(summary(tdf), caption = tabcap("Data summaries"))

```


Variables organization
=======================


Variables                       Values                  Treatment
------------------------        ---------------         ------------
__Demographic variables__
    * Sex                       female ; male           Make factor
    * Age                       numeric
__Family context__
    * SibSp                     numeric                 Combined
    * Parch                     numeric                 Combined
    * Famly                     numeric                 new: SibSp + Parch
__Relationship to ship__
    * Pclass                    1,2,3                   Make factor
    * ticket                    ticket number
    * Fare                      numeric
    * Cabin                     Cabin nbr               not used
    * Embark                    C = Cherbourg,          Factor
                                Q = Queenstown, 
                                S = Southampton
__Survival__              
    * survived                  0/1                     Make factor

```{r mantable1, echo=FALSE, results="asis"}
tabcap("variable organization")
```


Data modifications
==================

* Make `Survived` and `Pclass` factors, 
* Create `Famly` = `SibSp + Parch` 

```{r datamod}
sf <- factor(tdf$Survived, labels=c("No", "Yes"))
# verification
# head(tdf$Survived)
# head(sf2)

tdf$Survived <- sf

# makefactor class
tdf$Pclass <- factor(tdf$Pclass)

# combine SibSp + Parch
tdf$Famly <- tdf$SibSp + tdf$Parch

tdf$Embarked <-  factor(tdf$Embarked)

```




Data Analysis
==================

Passenger Identity
-------------------

```{r ident1, out.width="0.32\\textwidth", warning=FALSE, fig.cap="Passenger Identity"}
tbsex <- table(tdf$Sex)
tbsex
ggplot(tdf, aes(Sex)) + geom_bar()
```

```{r ident2, out.width="0.32\\textwidth", warning=FALSE, fig.cap="Passenger Identity contd."}
sumAge <- summary(tdf$Age)

nc <- nclass.Sturges(tdf$Age)
ggplot(tdf, aes(Age)) + geom_histogram(bins=nc)
ggplot(tdf, aes(Age)) + geom_histogram(aes(y=..density..),bins=nc) + facet_grid(Sex ~ .)
```

__Family context__

```{r family, out.width="0.32\\textwidth", fig.cap="Family members on board"}
ggplot(tdf) + geom_bar(aes(SibSp))

ggplot(tdf) + geom_bar(aes(Parch))

ggplot(tdf) + geom_bar(aes(Famly))
```

Passenger class
-------------------

```{r ship, fig.cap="Passenger Classes"}
pander(prop.table(table(tdf$Pclass)))
ggplot(tdf) + geom_bar(aes(Pclass, fill = Pclass, y = ..count../sum(..count..) ))
```

```{r family-class, out.width="0.32\\textwidth", fig.cap="Passenger demographics by class"}
ggplot(tdf, aes(Sex)) + 
        geom_bar(aes(y = ..count../sum(..count..), fill = Pclass)) + 
        facet_grid(Pclass ~ .) +
        labs(y = "Proportion")

nc <- nclass.FD(tdf$Age[!is.na(tdf$Age)])
ggplot(tdf, aes(Age)) + 
        geom_histogram(aes(y=..density..,fill=Pclass),bins=nc) +
        facet_grid(Pclass ~ .)

ggplot(tdf, aes(Age)) + 
        geom_histogram(aes(y=..density..,fill=Pclass),bins=nc) +
        facet_grid(Pclass ~ Sex)


# ggplot(tdf) + geom_bar(aes(SibSp, y = ..prop..)) + facet_grid(Pclass ~ .)
# ggplot(tdf) + geom_bar(aes(Parch)) + facet_grid(Pclass ~ .)

ggplot(tdf) + 
        geom_bar(aes(Famly, y = ..prop.., fill=Pclass)) +
        facet_grid(Pclass ~ .)

ggplot(tdf) + 
        geom_bar(aes(Famly, y = ..prop.., fill=Pclass)) +
        facet_grid(Pclass ~ Sex)
```

Compare with reference: \@ref(fig:ship)

```{r fare}

# ncf <- nclass.FD(tdf$Fare) 
ncf <- 50
ggplot(tdf) + geom_histogram(aes(Fare), bins = ncf)
ggplot(tdf) + geom_histogram(aes(Fare, y=..density.., fill=Pclass), bins = ncf) + 
        facet_grid(Pclass ~.) 
```

Survival
--------

## overall

```{r survival}
pander(prop.table(table(tdf$Survived)), digits=2)

ggplot(tdf) + geom_bar(aes(Survived, y = ..prop.., group = 1))
```

## By category

```{r survival-by}
# by sex
pander(prop.table(table(tdf$Survived, tdf$Sex),
                  margin = 2),
       digits=2)

# by class
pander(prop.table(table(tdf$Survived, tdf$Pclass),
                  margin = 2),
       digits=2)

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., group = Sex)) + 
        facet_grid(Sex ~.)

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., group = Pclass)) + 
        facet_grid(Pclass ~.) 


```

Class and sex

```{r survival-by2, fig.cap="Survival by Sex and Class"}

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Sex)
```

Two more lines, for just a try: 
Citing references: 

1. a figure: figure \@ref(fig:ship)
2. and tables: data head table (table \@ref(tab:dataget)) and summary table (table \@ref(tab:datasum)) and table \@ref(tab:mantable1))



** Bye-Bye **



