---
title: 'Titanic - Kaggle - Some exploration'
subtitle: "Explore for classroom use"
author:  "Bruno Fischer Colonimos"
date: "`r format(Sys.Date(), '%d %B %Y')`"
# header-includes: \usepackage{graphicx} # for unmodded template
header-includes: \usepackage{float} # for fig.pos = 'H'
fontsize: 12pt
urlcolor: "blue"   ## external
linkcolor: "red" ## internal
documentclass: article
classoption: a4paper
geometry: "left=1.5cm,right=1.5cm,top=1.5cm,bottom=2cm,footskip=1cm"
footer: \thetitle \ \textemdash \ \thepage ## verify this (theTitle == thetitle)
output:
  # bookdown::pdf_document2:
  pdf_document:
    # template: test_latex_temp
    template: latex_temp2.latex
    highlight: monochrome
    number_sections: yes
    toc: yes
    toc_depth: 4
    # keep_tex: true
    fig_caption: yes
    keep_tex: yes
    
  # html_document:
  #   number_sections: yes
  #   toc: yes
  #   toc_depth: 4
  #   fig_caption: yes
---



***************************


<!-- Some Inline HTML (CSS) , formatting tables -->

<style>

table {
    width: 100%;
    border-top-style: solid;
    border-top-width: medium;
    border-top-color: initial;
    border-bottom-style: solid;
    border-bottom-width: medium;
    border-bottom-color: initial;
}

</style>

<!-- End Inline HTML -->



Preliminary code
================

```{r setup0, include=FALSE}
# Later
# =======
# knitr::opts_chunk$set(echo = TRUE, fig.keep="all" ,fig.width=4, fig.asp= 0.75 , fig.align="center", fig.keep ="all", fig.show="hold", out.width="0.48\\textwidth")

# , fig.align="center"
# not for HTML : out.width="0.48\\textwidth"
```

## Libraries and auxiliary code (install before running)

(Not fully echoed here)

### code config parameters
```{r, echo = TRUE}
# adjust Gggplot2 theme and color palette
bwtheme <- TRUE
specialpalette <- FALSE
showarning <- FALSE
datmis <- "Ukn"
```

### Required packages (install before running)

"caret", "ggplot2", pander


```{r libs_prep, echo =FALSE, warning=FALSE, message=FALSE, results='hide'}

library(knitr)
library(pander)
library(caret)
library(ggplot2)
library(dplyr)


# Source preparatory code "preparation.R"
# ========================

source("preparation.R")

# Set aliases

knitr::set_alias(w = "fig.width", a = "fig.asp")

```

```{r setup, echo =FALSE}
# setup chunk : default options
knitr::opts_chunk$set(echo = FALSE, fig.width = w.13, fig.asp = a.11 , fig.pos = 'H', fig.align = "center", fig.show = "hold", warning = showarning)
```

Data
=====

Get it
-------

The data is downloaded from Kaggle (https://www.kaggle.com/c/titanic) and saved. It is loaded here from disk.

```{r dataget}
# read
datadir <-  "data"
fname <-  "train.csv"
tdf <- read.csv(file.path(datadir, fname), stringsAsFactors = FALSE)

```
 
 * The data head is shown in table `r .ref("tab:datafirst")`

```{r datafirst}
pander(head(tdf), caption = tabcap("A glimpse of the data"))     
```



Variables organization and work to do  {#variables}
---------------------------------------------------


Variables                 type        Values                Treatment
------------------------  --------    -------------------   ------------------
__Demographic variables__
    * Sex                  String       female ; male       Make factor
    * Age                  numeric                          many NA's
    * Agestat              factor      enfant/ado/adulte    new:
__Family context__
    * SibSp                numeric                          Combined ->
    * Parch                numeric                          Combined ->
    * Famly                numeric                          new: SibSp + Parch
__Relationship to ship__
    * Pclass                             1,2,3              Make factor
    * ticket                            ticket number       not used
    * Fare                 numeric
    * Cabin                             Cabin nbr           modified -> Deck
    * Deck                 String                           new (from Cabin)
    * Embark               String       C = Cherbourg,       make Factor   
                                        Q = Queenstown, 
                                        S = Southampton
__Survival__              
    * survived             binary        0/1                Make factor

:`r tabcap(chunklabel = "dataorg", "Data organization")`



* Quick dataframe summary: table `r .ref("tab:datasum")`
```{r datasum}
pander(summary(tdf), caption = tabcap("Data summaries"))
```



Data modifications {#modifs}
----------------------------

* Make `Survived` , `Pclass` and `Embark` factors, 
* Create `Famly` = `SibSp + Parch` 
* Create  `Sex.Pclass`
* Substitute missing values with `r datmis` in variable `Embarked`
* added variables
    * `Agestat` = age status : "child", "teen", "adult" (cutoff ages = 12, 18)
    * `Title` = civility.
    * `Letticket` = Ticket number begins with letters (yes = 1, no = 0)
    * `Hascabin` = is the cabin number known? (yes = 1, no = 0)
    * `Deck`= if the cabin is known, the first letter is the Deck (T, A, B, C...), otherwise  "`r datmis`" 
    

```{r datamod, warning=FALSE}
source("Titanic_preprocessing.R")
# missing values before modifications
pander(t(numna_before), caption = "Before modifications, Number of missing values")

```




Data Analysis {#analysis}
==================

Passenger Identity {#demographic}
-------------------

### Sex and Age {#sexage}

* Gender

```{r ident1, warning=FALSE, fig.cap="Gender distribution"}

# opts.labels("f13"), fails
tbsex2 <- local({
        tbsex <- table(tdf$Sex)
        ptbsex <- prop.table(tbsex)
        tbsex2 <- rbind(tbsex, ptbsex)
        row.names(tbsex2) <- c( "Frequency", "Rel.Frequency")
        tbsex2
})

pander(tbsex2, digits = 3, caption = tabcap("Gender distribution"))

ggplot(tdf, aes(Sex)) + geom_bar(aes(y= ..prop.., group = 1))
```

Figure `r .ref("fig:ident1")` shows the gender distribution

```{r ident2, warning=FALSE, fig.cap="Age and Gender"}
sumAge <- summary(tdf$Age)
```

* Age : 
`r pander(sumAge, caption = "Age distribution")`

* Gender and age

```{r ident3,  warning=FALSE, fig.cap="Age and Gender"}
# nc <- nclass.Sturges(tdf$Age)
nc <- nclass.FD(tdf$Age[!is.na(tdf$Age)])

ggnolegend <- theme(legend.position = "none")

ggplot(tdf, aes(Age)) + geom_histogram(bins=nc)

ggplot(tdf, aes(Age)) + geom_histogram(aes(fill = Pclass), bins=nc) + 
        facet_grid(Pclass ~ .) + 
        labs(title = "Age by Class") + ggnolegend

ggplot(tdf, aes(Age)) + geom_histogram(aes(y=..density..),bins=nc) + 
        facet_grid(Sex ~ .)

ggplot(tdf, aes(Age)) + geom_histogram(aes(y=..density.., fill = Pclass),bins=nc) + 
        facet_grid(Pclass ~ Sex) +
        labs(title = "Age by Class + Sex") + ggnolegend

ggplot(tdf, aes(x=Sex, y=Age)) + geom_boxplot() +
        scale_x_discrete(limits = rev(levels(tdf$Sex))) +
        coord_flip() + scale_y_continuous(breaks = seq(0,90, by=10))

ggplot(tdf, aes(x=Pclass, y=Age, fill = Pclass)) + geom_boxplot() +
        scale_x_discrete(limits = rev(levels(factor(tdf$Pclass)))) +
        coord_flip() + scale_y_continuous(breaks = seq(0,90, by=10)) + 
        # facet_grid( Pclass ~ .) 
        labs(title = "Age By Class")+ ggnolegend

ggplot(tdf, aes(x=Sex, y=Age)) + geom_boxplot() +
        scale_x_discrete(limits = rev(levels(tdf$Sex))) +
        coord_flip() + scale_y_continuous(breaks = seq(0,90, by=10)) + 
        facet_grid( Pclass ~ .) + 
        labs(title = "By Sex and Class") + ggnolegend
```


Figure `r .ref("fig:", "ident3")` shows the age distributions of both genders


* Age Status :
```{r Agestatus}
ggplot(tdf, aes(Agestat)) + geom_bar()

ggplot(tdf, aes(Agestat)) + 
        geom_bar(aes(y = ..prop.., group = Sex, fill = Sex)) + 
        facet_grid(Sex ~ .) + 
        theme(legend.position = "none")

ggplot(tdf, aes(Agestat)) + 
        # geom_bar() +
        geom_bar(aes(y = ..prop.., group = Pclass)) +
        facet_grid(. ~ Pclass)+
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5), legend.position = "none")

```
```{r Agestat2, w = w.11, fig.cap= "Age status by sex and class"}
ggplot(tdf, aes(Agestat)) + 
        geom_bar(aes(y = ..prop.., group = Pclass)) + 
        facet_grid(Sex ~ Pclass) +
        theme(axis.text.x = element_text(angle = 90))
```




Family members on board {#fmly}
---------------


```{r family,  fig.cap="Family members on board", warning=FALSE}

pander(cor(tdf[ , c("SibSp","Parch","Famly")]), caption = tabcap("Family members: correlation of the variables"))
```

```{r family2, fig.cap="Family members on board", warning=FALSE}
ggplot(tdf) + geom_bar(aes(SibSp))
ggplot(tdf) + geom_bar(aes(Parch))
ggplot(tdf) + geom_bar(aes(Famly))
```

Figure `r .ref("fig:", "family2")` shows the Family size variablesand their correlation. Famly seems to convey most of the information.


Passenger class {#pclass}
-------------------

figure `r .ref("fig:", "ship")` Shows that the third class accounts for about 55% of the passengers. 

```{r ship,  fig.cap="Passenger Classes Distribution", w = w.13, a = a.13}

pander(local({
        tb <- rbind(table(tdf$Pclass),
                    prop.table(table(tdf$Pclass)))
        
        row.names(tb) = c("frequency", "rel.frequency")
        tb}), 
       caption = tabcap("Passenger Class (Pclass)"))



ggplot(tdf) + geom_bar(aes(Pclass, fill = Pclass, group=1, y = ..prop.. )) +
        labs(title = "55% of the passengers are 3rd class")
```

```{r family_class, warning=FALSE , fig.cap="Passenger demographics by class"}

ggplot(tdf, aes(Sex)) + 
        geom_bar(aes(y = ..prop.., fill = Pclass, group=Pclass)) + 
        facet_grid(Pclass ~ .) +
        labs(y = "Proportion") +
        theme(legend.position = "none")

nc <- nclass.FD(tdf$Age[!is.na(tdf$Age)])
ggplot(tdf, aes(Age)) + 
        geom_histogram(aes(y=..density..,fill=Pclass),bins=nc) +
        facet_grid(Pclass ~ .) +
        theme(legend.position = "none")


ggplot(tdf, aes(Age)) + 
        geom_histogram(aes(y=..density..,fill=Pclass),bins=nc) +
        facet_grid(Pclass ~ Sex) +
        theme(legend.position = "none")



# ggplot(tdf) + geom_bar(aes(SibSp, y = ..prop..)) + facet_grid(Pclass ~ .)
# ggplot(tdf) + geom_bar(aes(Parch)) + facet_grid(Pclass ~ .)

ggplot(tdf) + 
        geom_bar(aes(Famly, y = ..prop.., fill=Pclass)) +
        facet_grid(Pclass ~ .)+ 
        scale_x_continuous(breaks=seq(0,10,by=2)) +
        theme(legend.position = "none")


ggplot(tdf) + 
        geom_bar(aes(Famly, y = ..prop.., fill=Pclass, group=Sex.Pclass)) +
        facet_grid(Pclass ~ Sex) + 
        scale_x_continuous(breaks=seq(0,10,by=2)) +
        theme(legend.position = "none")

```

Figure `r .ref("fig:", "family_class")`. Compare with reference : figure `r .ref("fig:ship")`

```{r Pclass_by_Famly, w = w.11, a = a.21}
ggplot(data = tdf, mapping = aes(Pclass,  fill= Pclass)) + 
        geom_bar() +
                facet_grid(factor(Famly) ~ .)

```


Deck
----

```{r Deck, fig.cap= "Distribution of decks"}
ggplot(tdf) +
        geom_bar(aes(Deck, y=..prop.., group = 1))

ggplot(tdf[tdf$Deck != datmis,] ) +
        geom_bar(aes(Deck, y=..prop.., group = 1))

ggplot(tdf[tdf$Deck != datmis,] ) +
        geom_bar(aes(Deck, y=..prop.., fill = Pclass, group = Pclass)) + 
        facet_grid(Pclass ~ .) + theme(legend.position = "none")        
 
```
Figure `r .ref("fig:", "Deck") shows a very partial distribution of decks: the deck of nearly 80% of the passengers is unknown. However, the third graph reveals the strong link between the passencger class and the deck.

Embarkation point
-----------------
The passengers could embark at Southampton (S) , Cherbourg or Cobh, alias Queenstown (Q) 

```{r Embarked, a= a.43, fig.cap="Embarkation port (C = Cherbourg, Q = Queenstown, S = Southampton)"}

tdf$Embarked <- factor(tdf$Embarked, levels = c("S", "C", "Q"))
temb <- table(tdf$Embarked)
# proportion table
ptemb <- prop.table(temb)
df <- rbind(frequency = temb , Rfreq = ptemb) 
pander(df, digits=2, caption = "Embarkation port")



# na_emb <- sum(is.na(tdf$Embarked)) #  2 passengers have unknown embarkation port
# 
# tdfx <- tdf[complete.cases(tdf), ] 

tdfx <- tdf 

ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., group=1))

ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., group=Sex)) + facet_grid(.~Sex) +
        theme(legend.position = "none")

ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., fill = Pclass,  group=Pclass)) + 
        facet_grid(Pclass~.) +
        theme(legend.position = "none")

ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., fill = Pclass, group=Pclass)) +
        facet_grid(Pclass~Sex) +
        theme(legend.position = "none")


```

```{r bydeckv, w = w.12, a = a.21}
ggplot(tdfx) + 
        geom_bar(aes(Embarked, y=..prop.., fill = Pclass)) +
        facet_grid(Deck~Sex) +
        theme(legend.position = "none") +
        labs(title = "Broken down by deck")
```


Figure `r .ref("fig:", "Embarked")` shows that Southampton was the major embarkation point (`r  100 * df["Rfreq", "S"]`% of the passengers). However, this is not as true for the women, particularly for the women with a first-class ticket.( only about 50% of them embarked at Southampton) 


The Fare
--------

```{r fare, w=w.12,a=a.11, fig.cap="Fare by class and Sex"}

# ncf <- nclass.FD(tdf$Fare) 
ncf <- 50
ggplot(tdf) + geom_histogram(aes(Fare), bins = ncf)
# ggplot(tdf) + geom_histogram(aes(Fare, y=..density.., fill=Pclass), bins = ncf) + 
#         facet_grid(Pclass ~.) 

ggplot(tdf) + geom_boxplot(aes(Pclass,Fare, fill = Pclass), color = "grey50") + 
        scale_x_discrete(limits = rev(levels(tdf$Pclass))) +
        coord_flip() +
        facet_grid(Sex ~.) 
```

```{r faredeck, w= w.11, fig.asp = 2 , fig.cap = "fare by deck and class"}

# unique(tdf$Deck)
local( {
        tdf1 <- tdf
        # tdf1$Deck <- reorder(tdf1$Deck, tdf1$Fare, median)
        tdf1$Deck <- factor(tdf$Deck)
        ggplot(tdf1) + 
                geom_boxplot(aes(Deck, Fare), outlier.alpha = 0.2) +
                coord_flip() +
                facet_grid( Pclass ~ ., switch = "y")
}
)
```




Survival
--------

### overall

```{r survival, fig.cap="Overall survival"}
pander(prop.table(table(tdf$Survived)), digits=2, caption = "Suvival distributiond")

ggplot(tdf) + geom_bar(aes(Survived, y = ..prop.., group = 1))
```

### By category

```{r survival-by, fig.cap="Survival by sex or Passenger Class"}
# by sex
pander(prop.table(table(tdf$Survived, tdf$Sex),
                  margin = 2),
       digits=2,
       caption = "Survival by Sex")

# by class
pander(prop.table(table(tdf$Survived, tdf$Pclass),
                  margin = 2),
       digits=2,
       caption = "Survival by Passenger class")

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., group = Sex)) + 
        facet_grid(Sex ~.)

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., group = Pclass)) + 
        facet_grid(Pclass ~.) 


```

* Class and sex

```{r survival-by2, fig.cap="Survival by Sex and Class"}

ggnolegend <- theme(legend.position = "none")

ggplot(tdf) +
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass)) + 
        facet_grid(Pclass ~ Sex) + ggnolegend
        
```


* Class , sex, Port of embarkation

```{r survival-by3, fig.cap="Survival by Sex and Class / Embarked"}

ggplot(tdfx[tdfx$Embarked == "S", ]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Sex)+
        labs(title = "Emb. Southampton") + 
        theme(legend.position = "none")



ggplot(tdfx[tdfx$Embarked == "C", ]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Sex)+
        labs(title = "Emb. Cherbourg")+
        theme(legend.position = "none")


ggplot(tdfx[tdfx$Embarked == "Q", ]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Sex)+
        labs(title = "Emb. Queenstown")+
        theme(legend.position = "none")



```

Class and Age see figure `r .ref("fig:survival-by4")`

```{r survival-by4, fig.cap="Age by Survival  and Class", warning=FALSE}

ggplot(tdf) + 
        geom_histogram(aes(Age, y= ..density.., fill = Pclass), bins = nc) +
        facet_grid(Pclass ~ Sex) +
        labs(title= "All") + 
        theme(legend.position = "none")
        

ggplot(tdf[tdf$Survived == "Yes", ]) + 
        geom_histogram(aes(Age, y= ..density.., fill = Pclass), bins = nc) +
        facet_grid(Pclass ~ Sex) +
        labs(title= "Survivors") + 
        theme(legend.position = "none")

ggplot(tdf[tdf$Survived == "No", ]) + 
        geom_histogram(aes(Age, y= ..density.., fill = Pclass), bins = nc) +
        facet_grid(Pclass ~ Sex) +
        labs(title= "Non-Survivors") + 
        theme(legend.position = "none")
```

* Class and Agestat, Agestat and sex

```{r survival-by5, fig.cap="Survival by age status, Class and sex"}

agesex_class <- table(tdf$Agestat, tdf$Sex, tdf$Pclass)
pander(agesex_class, caption = "Distribution of age-status vs sex and Pclass" )




ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Agestat) + 
        labs(title = "Sex = All") +
        theme(legend.position = "none")

ggplot(tdf[tdf$Sex == "female" ,]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Agestat) +
        labs(title = "Sex ='female'") + 
        theme(legend.position = "none")

ggplot(tdf[tdf$Sex == "male" ,]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Agestat)+
        labs(title = "Sex = 'male'") + 
        theme(legend.position = "none")

```


```{r survival_by6, w = w.11, fig.asp = 1.5, fig.cap="Survival by Deck, age and class"}

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Deck, fill = Pclass))+ 
        facet_grid(Deck ~ Agestat)+
        labs(title = "by Deck") # + theme(legend.position = "none")
```

```{r survival_by7, w = w.11, fig.asp = 1.5, fig.cap="Survival by Deck, Age and Sex"}

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Deck, fill = Sex))+ 
        facet_grid(Deck ~ Agestat)+
        labs(Title = "by Deck") # + theme(legend.position = "none")
```


