---
title: 'Titanic - Kaggle - Prediction'
subtitle: "Explore for classroom use"
author:  "Bruno Fischer Colonimos"
date: "`r format(Sys.Date(), '%d %B %Y')`"
# header-includes: \usepackage{graphicx} # for unmodded template
fontsize: 12pt
urlcolor: "blue"   ## external
linkcolor: "red" ## internal
documentclass: article
classoption: a4paper
geometry: "left=1.5cm,right=1.5cm,top=1.5cm,bottom=2cm,footskip=1cm"
footer: \thetitle \ \textemdash \ \thepage 
output:
  # bookdown::pdf_document2:
  pdf_document:
    # template: test_latex_temp
    template: latex_temp2.latex
    highlight: monochrome
    number_sections: yes
    toc: yes
    toc_depth: 4
    # keep_tex: true
    fig_caption: yes
    
  # html_document:
  #   number_sections: yes
  #   toc: yes
  #   toc_depth: 4
  #   fig_caption: yes
---

***************************


<!-- Some Inline HTML (CSS) , formatting tables -->

<style>

table {
    width: 100%;
    border-top-style: solid;
    border-top-width: medium;
    border-top-color: initial;
    border-bottom-style: solid;
    border-bottom-width: medium;
    border-bottom-color: initial;
}

</style>

<!-- End Inline HTML -->



Preliminary code
================

```{r globalopts}
# Global options
globalseed <- 9876 # unique seed for reproducibility
percenttraining <- 0.75 # percentage of 'trainingset' to put in truetrainingset
nfolds <- 5 # k = number of folds fork-folds crossvalidation
```


## Required packages (install before running)

"caret", "ggplot2", pander, doParallel    



## Libraries and auxiliary code (install before running)

(Not fully echoed here)

```{r, echo = FALSE}
# adjust Gggplot2 theme and color palette
bwtheme <- TRUE
specialpalette <- TRUE # controls the colour of the plots
showarning <- TRUE # show warnings (deleted empty rows...)
datmis <- "Ukn"
```


```{r libs, echo =FALSE, warning=FALSE, message=FALSE, results='hide'}
# libraries
library(pander)
library(ggplot2)
library(dplyr)
library(knitr, quietly = TRUE)
library(caret, quietly = TRUE )
library(parallel, quietly = TRUE)
library(doParallel, quietly = TRUE)
library(rpart, quietly = TRUE)
library(randomForest, quietly = TRUE)
library(MASS, quietly = TRUE)
library(klaR, quietly = TRUE)
library(fastAdaboost, quietly = TRUE)
library(e1071, quietly = TRUE)

# library(party) #trees
# library(ISLR) # example
# 
# library(ElemStatLearn)

# Source preparatory code
# ========================

source("preparation.R")

# Set aliases
knitr::set_alias(w = "fig.width", a = "fig.asp")

# default options
knitr::opts_chunk$set(echo = FALSE, fig.width = w.13, fig.asp = a.11 , fig.pos = 'H', fig.align = "center", fig.show = "hold")


```

Data
=====

Get it
-------

```{r dataget}
# read
datadir <-  "data"
fname <-  "train.csv"
tdf <- read.csv(file.path(datadir, fname), stringsAsFactors = FALSE)

```

Variables organization {#variables}
-----------------------


Variables                 type        Values                Treatment
------------------------  --------    -------------------   ------------------
__Demographic variables__
    * Sex                  String       female ; male       Make factor
    * Age                  numeric
    * Agestat              factor      enfant/ado/adulte    new:
__Family context__
    * SibSp                numeric                          Combined ->
    * Parch                numeric                          Combined ->
    * Famly                numeric                          new: SibSp + Parch
__Relationship to ship__
    * Pclass                             1,2,3              Make factor
    * ticket                            ticket number       not used
    * Fare                 numeric
    * Cabin                             Cabin nbr           not used
    * Embark               Factor       C = Cherbourg,          
                                        Q = Queenstown, 
                                        S = Southampton
__Survival__              
    * survived             binary        0/1                Make factor

:`r tabcap(chunklabel = "dataorg", "Data organization")`


Data modifications {#modifs}
----------------------------

* Make `Survived` , `Pclass` and `Embark`   factors, 
* Create `Famly` = `SibSp + Parch` 


```{r datamod, warning=FALSE}
sf <- factor(tdf$Survived, labels=c("No", "Yes"))
tdf$Survived <- sf

# makefactor class
tdf$Pclass <- factor(tdf$Pclass)

# combine SibSp + Parch
tdf$Famly <- tdf$SibSp + tdf$Parch

# Embarked
tdf$Embarked[which(tdf$Embarked == "")] <- datmis
tdf$Embarked <-  factor(tdf$Embarked)
# levels(tdf$Embarked)
# levels(tdf$Embarked)

# ajout de variables
tdf <- tdf %>%
        mutate(sex.class = paste(Sex, Pclass, sep  = "."))

# age status
tdf$agestat <- cut(tdf$Age, breaks=c(0,12,18,100), labels=c("child", "teen", "adult"))

tdf$agestat <- as.character(tdf$agestat)
tdf$agestat <- ifelse(is.na(tdf$agestat), datmis, tdf$agestat)
tdf$agestat <- factor(tdf$agestat)
# levels(tdf$agestat)
# table(tdf$agestat)
```

### Cleaning: transforming covariates, removing superfluous variables

```{r dataclean1, warning= FALSE}
# let_ticket = are letters included in the ticket number
tdf$letticket <- as.numeric(substr(tdf$Ticket,start = 1, stop = 1 ) %in% LETTERS)

# hascabin is the cabin number nkown?
tdf$hascabin <- as.numeric(tdf$Cabin != "")
# sum(tdf$hascabin)


# Deck (from https://triangleinequality.wordpress.com/2013/09/08/basic-feature-engineering-with-the-titanic-data/)

#This is going be very similar, we have a ‘Cabin’ column not doing much, only
#1st class passengers have cabins, the rest are ‘Unknown’. A cabin number looks
#like ‘C123’. The letter refers to the deck, and so we’re going to extract these
#just like the titles. Turning cabin number into Deck
#
# cabin_list <- c('A', 'B', 'C', 'D', 'E', 'F', 'T', 'G', 'Unknown')
# df[,'Deck']=df[,'Cabin'].map(lambda x: substrings_in_string(x, cabin_list))

tdf$Deck <- factor(ifelse( tdf$Cabin != "", 
                           substr(tdf$Cabin,start = 1,stop = 1), 
                           datmis))
tdf$Deck <- relevel(tdf$Deck, ref = "T")
# levels(tdf$Deck)


# Representing the factors with dummies
dummies <- dummyVars(Survived ~ ., data = tdf[c("Survived", "Pclass", "Sex", "Embarked", "Deck" )])
dumvars <- predict(dummies, newdata = tdf[c("Survived", "Pclass", "Sex", "Embarked", "Deck" )])
# dumvars <- dumvars[, c(1, 2, 4, 6, 7, 8, )]
dumvars <- as.data.frame(dumvars)
dumvars <- dumvars[-c(3, 5, 9, 18)]
# colnames(dumvars)
# str(dumvars)

# add dummies to the data
tdf <- cbind(tdf, dumvars)
# names(tdf)



# tdf1 <- tdf %>%
#         dplyr::select(-c(PassengerId, Name, Ticket, Cabin, sex.class))
# tdf1 <- tdf1[complete.cases(tdf1), ]


tdf2 <- tdf[, c("Survived", 
                "Age", "SibSp" ,"Parch","Fare", 
                # "Cabin", 
                # "Embarked",   
                "letticket",  "hascabin", 
                "Pclass.1", "Pclass.2", 
                 "Sexfemale",  "Embarked.C", "Embarked.Q", "Embarked.S",
                "Deck.T", "Deck.A", "Deck.B", "Deck.C", "Deck.D", "Deck.E", "Deck.F", "Deck.G" )]



# tdf2 <- tdf %>%
#         dplyr::select(-c(PassengerId, Name, Ticket, Cabin, sex.class, agestat, Famly)) %>%
#         dplyr::select(-c(Pclass, Sex)) # remove factors, keep dummies



# What to do with missing values ???

tdf2 <- tdf2[complete.cases(tdf2), ]
# tdf2 <- tdf2



nzv <- nearZeroVar(tdf2)

tdf2 <- tdf2[-nzv]

# **** PROBLEM: Later, we have an error : variables are collinear !! ***(OLD)

# cormat <- cor(tdf2[,-1])
# 
# highlyCorDescr <- findCorrelation(cormat, cutoff = .90)
# 
# hinames <- names(tdf2)[highlyCorDescr]
# # 
# # head(tdf[, hinames])
# # 
# # cor(data.frame(#tdf$hascabin,
# #                tdf$letticket,
# #                tdf$Pclass.1, 
# #                tdf$Pclass.2))

# 
# # findLinearCombos(tdf2[-1])


```



### Splitting data

```{r datasplit}
set.seed(globalseed)
intrain <- createDataPartition(y=tdf2$Survived,
                               p= percenttraining,
                               list = FALSE)

trainingset <- tdf2[intrain,]
validationset <- tdf2[-intrain,]
```






Prediction attempts
====================

Tools
-----

```{r modelbuildtools}
# functions to record modeling attempts
recordmod <- local({
        recdf <- data.frame(
                Name = character(),
                Method = character(),
                Preprocess = character(),
                Best.tune = character(),
                Accuracy.CV = numeric(),
                Accuracy.validation = numeric(),
                Model.name = character(),
                stringsAsFactors = FALSE)
        reclist <- list(recdf)
        init <- function(i = 1){ # i = the record "track"
                recdf <- data.frame(
                        Name = character(),
                        Method = character(),
                        Preprocess = character(),
                        Best.tune = character(),
                        Accuracy.CV = numeric(),
                        Accuracy.validation = numeric(),
                        Model.name = character(),
                        stringsAsFactors = FALSE)
                reclist <- reclist
                reclist[[i]] <- recdf
                reclist <<- reclist # update list in parent env
        }
        new <- function(i = 1, # the record "track" = either an integer or a name (string)
                        Name ="",
                        Method = "" ,
                        Preprocess ="" ,
                        Best.tune ="",
                        Accuracy.CV = 0 ,
                        Accuracy.validation = 0,
                        Model.name = "") {
                reclist <- reclist
                recdf <- reclist[[i]]
                newrow <- data.frame(Name = Name,
                                     Method = Method,
                                     Preprocess = Preprocess,
                                     Best.tune = Best.tune,
                                     Accuracy.CV = Accuracy.CV ,
                                     Accuracy.validation = Accuracy.validation,
                                     Model.name = Model.name,
                                     stringsAsFactors = FALSE )
                recdf <- rbind(recdf, newrow)
                reclist[[i]] <- recdf # update
                reclist <<- reclist # update
        }
        get <- function(i = 1) reclist[[i]]
        # export funtions
        list(init = init,
             new=new,
             get=get)
})

# initialize tracks 1 and 2
recordmod$init(1)
recordmod$init(2)

```






Model Building
--------------

### KNN

```{r modelbuildknn, cache=TRUE, dependson="dataclean1"}
# _______________________________
# Knn
# 
# parallel processing
# cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
# registerDoParallel(cluster)


set.seed(globalseed)
modl <- modknn <- train(Survived ~ . , data = trainingset,
                        method = "knn",
                        # preProcess = "pca",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
# stopCluster(cluster)
# registerDoSEQ()
```


```{r modelbuildknnrec}
# record attempt
# modl
cm <- confusionMatrix(data = predict( modl , newdata=validationset, type="raw"),
                reference = validationset$Survived)

recordmod$new(Name = "KNN", 
              Method = modl$method, 
              Preprocess = "None",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modknn" )
# recordmod$get()
```

### KNN-PCA

```{r modelbuildknn_pca, cache=TRUE, dependson="dataclean1"}
# _______________________________
# Knn
# 
# parallel processing
# cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
# registerDoParallel(cluster)


set.seed(globalseed)
modl <- modknn_pca <- train(Survived ~ . , data = trainingset,
                        method = "knn",
                        preProcess = "pca",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
# stopCluster(cluster)
# registerDoSEQ()
```


```{r modelbuildknnrec_pca}
# record attempt
# modl
cm <- confusionMatrix(data = predict( modl , newdata=validationset, type="raw"),
                reference = validationset$Survived)

recordmod$new(Name = "KNN_PCA", 
              Method = modl$method, 
              Preprocess = "Pca",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modknn" )
# recordmod$get()
```


### Random Forest

```{r modelbuildrf, cache=TRUE, dependson="dataclean1"}
# _______________________________
# Random Forests
#  
# parallel processing
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)



set.seed(globalseed)
modl <- modrf <- train(Survived ~ . , data = trainingset,
                        method = "rf",
                        # preProcess = "pca",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
stopCluster(cluster)
registerDoSEQ()
```


```{r modelbuildrfrec}
# record attempt
# modl
cm <- confusionMatrix(data = predict( modl , newdata=validationset, type="raw"),
                reference = validationset$Survived)

recordmod$new(Name = "RF", 
              Method = modl$method, 
              Preprocess = "None",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modrf" )

```



### Random Forest + PCA

```{r modelbuildrf_pca, cache=TRUE, dependson="dataclean1"}
# _______________________________
# Random Forests - PCA
#  
# parallel processing
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)



set.seed(globalseed)
modl <- modrf_pca <- train(Survived ~ . , data = trainingset,
                        method = "rf",
                        preProcess = "pca",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
stopCluster(cluster)
registerDoSEQ()
```


```{r modelbuildrfrec_pca}
# record attempt
# modl
cm <- confusionMatrix(data = predict( modl , newdata=validationset, type="raw"),
                reference = validationset$Survived)

recordmod$new(Name = "RF_PCA", 
              Method = modl$method, 
              Preprocess = "PCA",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modrf" )

```


### Glm

```{r modelbuildglm, cache=TRUE, dependson="dataclean1"}
# _______________________________
# glm
# 
# parallel processing
# cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
# registerDoParallel(cluster)


set.seed(globalseed)
modl <- modglm <- train(Survived ~ . , data = trainingset,
                        method = "glm",
                       family = "binomial",
                        # preProcess = "pca",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
# stopCluster(cluster)
# registerDoSEQ()
```


```{r modelbuildglmrec}
# record attempt
cm <- confusionMatrix(data = predict( modl , newdata=validationset, type="raw"),
                reference = validationset$Survived)

recordmod$new(Name = "GLM", 
              Method = modl$method, 
              Preprocess = "None",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modglm" )

# recordmod$get()
```



### Glm_pca

```{r modelbuildglm_pca, cache=TRUE, dependson="dataclean1"}
# _______________________________
# glm
# 
# parallel processing
# cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
# registerDoParallel(cluster)


set.seed(globalseed)
modl <- modglm_pca <- train(Survived ~ . , data = trainingset,
                        method = "glm",
                       family = "binomial",
                        preProcess = "pca",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
# stopCluster(cluster)
# registerDoSEQ()
```


```{r modelbuildglm_pca_rec}
# record attempt
cm <- confusionMatrix(data = predict( modl , newdata=validationset, type="raw"),
                reference = validationset$Survived)

recordmod$new(Name = "GLM_PCA", 
              Method = modl$method, 
              Preprocess = "pca",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modglm_pca" )

# recordmod$get()
```


### Glm with interaction

```{r modelbuildglm_int, cache=TRUE, dependson="dataclean1"}
# _______________________________
# glm
# 
# parallel processing
# cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
# registerDoParallel(cluster)


set.seed(globalseed)
modl <- modglm_int <- train(Survived ~ Age + SibSp + Parch + Fare + letticket + hascabin + Pclass.1 + Pclass.2 + Sex.male + Embarked.C + Embarked.Q + Age * Pclass.1 + Age * Pclass.2 + Age * Sex.male * Pclass.1 + + Age * Sex.male * Pclass.2,
                            data = trainingset,
                            method = "glm",
                            family = "binomial",
                            # preProcess = "pca",
                            trControl = trainControl(method = "cv",
                                                     number = nfolds))

# stop parallel proc.
# stopCluster(cluster)
# registerDoSEQ()
```


```{r modelbuildglm_int_rec}
# record attempt
cm <- confusionMatrix(data = predict( modl , 
                                      newdata=validationset, 
                                      type="raw"),
                      reference = validationset$Survived)

recordmod$new(Name = "GLM_INT", 
              Method = modl$method, 
              Preprocess = "None",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modglm_int" )

# recordmod$get()
```

### LDA

```{r modelbuildlda, cache=TRUE, dependson="dataclean1"}
# _______________________________
# lda
# 
# parallel processing
# cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
# registerDoParallel(cluster)


set.seed(globalseed)
modl <- modlda <- train(Survived ~ . , data = trainingset,
                        method = "lda",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
# stopCluster(cluster)
# registerDoSEQ()
```


```{r modelbuildldarec}
# record attempt
cm <- confusionMatrix(data = predict( modl , newdata=validationset, type="raw"),
                reference = validationset$Survived)

recordmod$new(Name = "LDA", 
              Method = modl$method, 
              Preprocess = "None",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modlda" )

# recordmod$get()
```

### LDA-PCA

```{r modelbuildlda_pca, cache=TRUE, dependson="dataclean1"}
# _______________________________
# lda
# 
# parallel processing
# cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
# registerDoParallel(cluster)


set.seed(globalseed)
modl <- modlda_pca <- train(Survived ~ . , data = trainingset,
                        method = "lda",
                        preProcess = "pca",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
# stopCluster(cluster)
# registerDoSEQ()
```


```{r modelbuildlda_pca_rec}
# record attempt
cm <- confusionMatrix(data = predict( modl , newdata=validationset, type="raw"),
                reference = validationset$Survived)

recordmod$new(Name = "LDA_PCA", 
              Method = modl$method, 
              Preprocess = "pca",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modlda" )

# recordmod$get()
```

### QDA

```{r modelbuildqda, cache=TRUE, dependson="dataclean1"}
# _______________________________
# qda
# 
# parallel processing
# cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
# registerDoParallel(cluster)


set.seed(globalseed)
modl <- modqda <- train(Survived ~ . , data = trainingset,
                        method = "qda",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
# stopCluster(cluster)
# registerDoSEQ()
```


```{r modelbuildqdarec}
# record attempt
cm <- confusionMatrix(data = predict( modl , newdata=validationset, type="raw"),
                reference = validationset$Survived)

recordmod$new(Name = "QDA", 
              Method = modl$method, 
              Preprocess = "None",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modlda_pca" )

# recordmod$get()
```


### Naive Bayes

```{r modelbuildnb, cache=TRUE, dependson="dataclean1", warning=FALSE}
# _______________________________
# nb = naive Bayes
# 
# parallel processing
# cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
# registerDoParallel(cluster)


set.seed(globalseed)
modl <- modnb <- train(Survived ~ . , data = trainingset,
                        method = "nb",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
# stopCluster(cluster)
# registerDoSEQ()
```


```{r modelbuildnbrec, warning=FALSE}
# record attempt
cm <- confusionMatrix(data = predict( modl , newdata=validationset, type="raw"),
                reference = validationset$Survived)

recordmod$new(Name = "NB", 
              Method = modl$method, 
              Preprocess = "None",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modnb" )

# recordmod$get()
```


### Naive Bayes - PCA

```{r modelbuildnb_pca, cache=TRUE, dependson="dataclean1", warning=FALSE}
# _______________________________
# nb = naive Bayes - PCA
# 
# parallel processing
# cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
# registerDoParallel(cluster)


set.seed(globalseed)
modl <- modnb_pca <- train(Survived ~ . , data = trainingset,
                        method = "nb",
                        preProcess = "pca",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
# stopCluster(cluster)
# registerDoSEQ()
```


```{r modelbuildnbrec_pca, warning=FALSE}
# record attempt
cm <- confusionMatrix(data = predict( modl , newdata=validationset, type="raw"),
                reference = validationset$Survived)

recordmod$new(Name = "NB_PCA", 
              Method = modl$method, 
              Preprocess = "None",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modnb_pca" )

# recordmod$get()
```


Summary of models
---------------

```{r}
kable(recordmod$get(), row.names = FALSE, caption = "Base models accuracy")
```





Stacking models
---------------

### Base models to stack

```{r}
basemods <- data.frame(model= c(
        "modknn_pca", 
        "modrf",
        "modglm",
        "modlda_pca",
        "modqda",
        "modnb_pca")
)

```



### predictions from base models

```{r basepreds, warning=FALSE}
# trainingset
stk_train_df = data.frame(
        p_knn = predict(modknn_pca, newdata = trainingset, type="prob"),
        p_rf = predict(modrf, newdata = trainingset, type="prob"),
        p_glm = predict(modglm, newdata = trainingset, type="prob"),
        # p_glm_int = predict(modglm_int, newdata = trainingset, type="prob"),
        p_lda = predict(modlda_pca, newdata = trainingset, type="prob"),
        p_qda = predict(modqda, newdata = trainingset, type="prob"),
        p_nb = predict(modnb_pca, newdata = trainingset, type="prob"),
        Survived =trainingset$Survived
)

# validationset
stk_valid_df = data.frame(
        p_knn = predict(modknn, newdata = validationset, type="prob"),
        p_rf = predict(modrf, newdata = validationset, type="prob"),
        p_glm = predict(modglm, newdata = validationset, type="prob"),
        # p_glm_int = predict(modglm_int, newdata = validationset, type="prob"),
        p_lda = predict(modlda, newdata = validationset, type="prob"),
        p_qda = predict(modqda, newdata = validationset, type="prob"),
        p_nb = predict(modnb, newdata = validationset, type="prob"),
        Survived =validationset$Survived
)


```


### stacked predictions

#### Combined With RF

```{r modelbuild_stk_rf, cache=TRUE}
# stkrf Stacked preds with rf
# 
# parallel processing
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)


set.seed(globalseed)
modl <- modstkrf <- train(Survived ~ . , data = stk_train_df,
                        method = "rf",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
stopCluster(cluster)
registerDoSEQ()
```



```{r modelbuild_stk_rf_rec}
# record attempt
cm <- confusionMatrix(data = predict( modl , newdata=stk_valid_df, type="raw"),
                reference = validationset$Survived)

recordmod$new( i = 2,
              Name = "Stacked RF", 
              Method = modl$method, 
              Preprocess = "None",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modstkrf" )

```



#### Combined With adaboost

```{r modelbuild_stk_ada, cache=TRUE}
# stkrf Stacked preds with rf
# 
# parallel processing
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)


set.seed(globalseed)
modl <- modstk_ada <- train(Survived ~ . , data = stk_train_df,
                        method = "adaboost",
                        trControl = trainControl(method = "cv",
                                                 number = nfolds))

# stop parallel proc.
stopCluster(cluster)
registerDoSEQ()
```



```{r modelbuild_stk_ada_rec}
# record attempt
cm <- confusionMatrix(data = predict( modl , newdata=stk_valid_df, type="raw"),
                reference = validationset$Survived)

recordmod$new( i = 2,
              Name = "Stacked adaboost", 
              Method = modl$method, 
              Preprocess = "None",
              Best.tune = paste0(colnames(modl$bestTune)[1], " = ", modl$bestTune[1]),
              Accuracy.CV = round(max(modl$results$Accuracy),3),
              Accuracy.validation = round(cm$overal["Accuracy"],3),
              Model.name = "modstk_ada" )

```



```{r}
kable(recordmod$get(i=2), row.names = FALSE, caption = "Stacked RF")
```


