---
title: 'Titanic - Kaggle - Some exploration'
subtitle: "Explore for classroom use"
author:  "Bruno Fischer Colonimos"
date: "`r format(Sys.Date(), '%d %B %Y')`"
# header-includes: \usepackage{graphicx} # for unmodded template
fontsize: 12pt
urlcolor: "blue"   ## external
linkcolor: "red" ## internal
documentclass: article
classoption: a4paper
geometry: "left=1.5cm,right=1.5cm,top=1.5cm,bottom=2cm,footskip=1cm"
footer: \thetitle \ \textemdash \ \thepage 
output:
  # bookdown::pdf_document2:
  pdf_document:
    # template: test_latex_temp
    template: latex_temp2.latex
    highlight: monochrome
    number_sections: yes
    toc: yes
    toc_depth: 4
    # keep_tex: true
    fig_caption: yes
    
  # html_document:
  #   number_sections: yes
  #   toc: yes
  #   toc_depth: 4
---
----------------


<!-- Some Inline HTML (CSS) , formatting tables -->

<style>

table {
    width: 100%;
    border-top-style: solid;
    border-top-width: medium;
    border-top-color: initial;
    border-bottom-style: solid;
    border-bottom-width: medium;
    border-bottom-color: initial;
}

</style>

<!-- End Inline HTML -->



Preliminary code
================


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.keep="all" ,fig.width=4, fig.asp= 0.75 , fig.align="center", fig.keep ="all", fig.show="hold", out.width="0.48\\textwidth")

# , fig.align="center"
# not for HTML : out.width="0.48\\textwidth"
```


## Required packages (install before running)

"caret", "ggplot2", pander       



## Libraries and auxiliary code (install before running)

(Not fully echoed here)

```{r}
# adjust Gggplot2 theme and color palette
bwtheme <- TRUE
specialpalette <- TRUE # controls the colour of the plots
```


```{r libs, echo =FALSE, warning=FALSE, message=FALSE, results='hide'}
library(knitr)
library(pander)
library(caret)
library(ggplot2)
library(dplyr)
# library(party) #trees
# library(ISLR) # example
# 
# library(ElemStatLearn)
if (bwtheme) {
        theme_set(theme_bw())
}

if (specialpalette) {
        # palette color-blind-friendly: The palette with black (Cookbook for R,
        # http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/)
        cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73",
                        "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
        
        # modified BFC
        cbfPalette <- cbbPalette
        cbfPalette[3] <- cbbPalette[7]
        cbfPalette[6] <- cbbPalette[3]
        cbfPalette[8] <- cbbPalette[7]
        
        # * `set_default_scale` has been removed. If you want to change the default
        # scale for an aesthetic, just create a function called
        # `scale_aesthetic_continuous` or `scale_aesthetic_discrete` that returns the
        # scale that you want. 
        
        scale_colour_discrete <- function(...) {
                scale_colour_manual(values = cbbPalette, ...)
        }
        
        scale_fill_discrete <- function(...) {
                scale_fill_manual(values = cbbPalette, ...)
        }
        
        scale_colour_continuous <- function(...) {
                ggplot2::scale_colour_brewer("Greys", ...)
        }
        
}



# ** Some code *******
#######################################################
# Identify context: output type ***********************


# From Hadley's ggplot2 book:  Knowing conversion target 
is_latex <- function() {
        identical(knitr::opts_knit$get("rmarkdown.pandoc.to"), "latex")
}

# html same
is_html <- function() {
        identical(knitr::opts_knit$get("rmarkdown.pandoc.to"), "html")
}

pdfoutput <- is_latex() # tells if we are converting through latex, ie to pdf

# getcontext returns the doc conversion context
getcontext <- function(){
        if (is_latex()) {"latex" 
                   } else if (is_html()) {
                           "html"
                   } else {"other"} 
}


# get the code chunk name (for captionning)

chunkname <- function() {knitr::opts_current$get("label")}


#######################################################
# Captioning / cross-referencing ***********************


# Table captions
# tabcap <- function(caption,
#                    chunklabel=knitr::opts_current$get("label"),
#                    context=getcontext()) {
#         
#         switch(context,
#                latex = paste0("\\label{tab:", chunklabel, "}",
#                               caption),
#                html = paste0("<div id=tab:", chunklabel, ">",
#                              caption, "</div>"),
#                other = caption ) 
# }


tabcap <- function(caption,
                   chunklabel=knitr::opts_current$get("label"),
                   context=getcontext()) {
        
        switch(context,
               latex = paste0("\\label{tab:", chunklabel, "}",
                              caption),
               html = paste0("<a name=tab:", chunklabel, ">",
                             caption, "</a>"),
               other = caption ) 
}







# figure captions ?? ==> No, not needed with pdf_output
figcap <- function(caption, chunklabel=knitr::opts_current$get("label")) {
        # debug
        # thislabel <<- chunklabel
        paste0("\\label{fig:", chunklabel, "}", caption)
}

# tests

# tabcap("one", "thzchunk", context = "latex")
# tabcap("one", "thzchunk", context = "html")


# inserting a reference to a label, using the context

.ref <- function(prefix = NULL,
                 reflabel = "",
                 context = getcontext()) {
        switch(EXPR = context,
               latex = paste0("\\ref{",prefix, reflabel,"}"),
               html = paste0("<A HREF=\\#",
                              prefix, reflabel,
                              "\\>", reflabel , "</A>"), # reflabel here ?
               other = paste0("(#",prefix, reflabel ,")")
        )
}

# .ref <- function(prefix = NULL, 
#                  reflabel = "", 
#                  context = getcontext()) {
#         switch(EXPR = context,
#                latex = paste0("\\ref{",prefix, reflabel,"}"),
#                html = paste0("<a href= #", 
#                               prefix, reflabel,
#                               ">", reflabel , "</a>"), # reflabel here ?
#                other = paste0("(#",prefix, reflabel ,")")
#         )        
# }






# testit

# .ref(prefix= "tab:",reflabel = "onelab", context =  "latex" )
# .ref(prefix= "fig:", reflabel = "onelab", context = "html" )
# .ref(reflabel = "onelab", context = "html" )
# .ref(prefix= "fig:", reflabel = "onelab", context = "other" )


#######################################################
# figures dimensions parameters ***********************

#   aspect ratio
a.13 <- 1 / 3
a.12 <- 0.5
a.34 <- 0.75
a.11 <- 1
a.43 <- 1.33
a.21 <- 2

#  width
# == doc line width in inches * security coeff 0.98

doclinewidth <- (21 - 1.5 - 1.5) * 0.3937 * 0.98  # légère correction

w.11 <-  doclinewidth
w.34 <- 3 / 4 * doclinewidth
w.23 <- 2 / 3 * doclinewidth
w.12  <-  1 / 2 * doclinewidth
w.13 <- 1 / 3 * doclinewidth
w.14 <- 1 / 4 * doclinewidth * 0.99 # de nouveau: légère correction

# #   output width (only pdf)
# one3outwidth <- if (pdfoutput) {"0.32\\textwidth"} else {NULL}
# halfoutwidth <- if (pdfoutput) {"0.48\\textwidth"} else {NULL}
# two3outwidth <- if (pdfoutput) {"0.64\\textwidth"} else {NULL}
# fulloutwidth <- if (pdfoutput) {"\\textwidth"} else {NULL}

# configs pr?d?finies knitr::opts_template$set()

.codeR <- paste0( "knitr::opts_template$set(",
                  if (pdfoutput) {paste0(".f14 = list(fig.width = ", 
                                         w.14, 
                                         ", fig.asp = ", 
                                         a.12, "), ")
                          } else { # w.14 too narrow to render ok in html
                                  paste0(".f14 = list(fig.width = ", 
                                         w.13, 
                                         ", fig.asp = ", 
                                         a.34, "), ")
                                  
                          }  ,
                  "f13 = list(fig.width = ", w.13, ", fig.asp = ", a.34, "), ",
                  "f12 = list(fig.width = ", w.12, ", fig.asp = ", a.34, "), ",
                  "fnorm = list(fig.width = ", w.23, ", fig.asp = ", a.34, "), ",
                  "fnormsq = list(fig.width = ", w.23, ", fig.asp = ", a.11, "), ",
                  "fnormhigh = list(fig.width = ", w.23, ", fig.asp = ", a.43, "), ",
                  "fnormvhigh = list(fig.width = ", w.23, ", fig.asp = ", a.21, "), ",
                  "fulllow = list(fig.width = ", w.11, ", fig.asp = ", a.34, "),",
                  "full = list(fig.width = ", w.11, ", fig.asp = ", a.34, "),",
                  "fullsq = list(fig.width = ", w.11, ", fig.asp = ", a.11, "),",
                  "fullhigh = list(fig.width = ", w.11, ", fig.asp = ", a.43, "),",
                  "fullvhigh = list(fig.width = ", w.11, ", fig.asp = ", a.21, ")",
                  ")"
)


eval(parse(text = .codeR))  
              
              
# verif
# knitr::opts_template$get()
# knitr::opts_template$get(".f14")

# Set aliases
knitr::set_alias(w = "fig.width", a = "fig.asp")

# default options
knitr::opts_chunk$set(echo = FALSE, fig.width = w.23, fig.asp = a.34 , fig.pos = 'H', fig.align = "center", fig.show = "hold")


```

Data
=====

Get it
-------

```{r dataget}
# read
datadir <-  "data"
fname <-  "train.csv"
tdf <- read.csv(file.path(datadir, fname))

```
 
 * The data. Table `r .ref("tab:datafirst")`

```{r datafirst}
pander(head(tdf), caption = tabcap("A glimpse of the data"))     
```

* Quick dataframe summary: table `r .ref("tab:datasum")`
```{r datasum}
pander(summary(tdf), caption = tabcap("Data summaries"))
```


Variables organization {#variables}
-----------------------


Variables                       Values                  Treatment
------------------------        ---------------         ------------
__Demographic variables__
    * Sex                       female ; male           Make factor
    * Age                       numeric
__Family context__
    * SibSp                     numeric                 Combined
    * Parch                     numeric                 Combined
    * Famly                     numeric                 new: SibSp + Parch
__Relationship to ship__
    * Pclass                    1,2,3                   Make factor
    * ticket                    ticket number
    * Fare                      numeric
    * Cabin                     Cabin nbr               not used
    * Embark                    C = Cherbourg,          Factor
                                Q = Queenstown, 
                                S = Southampton
__Survival__              
    * survived                  0/1                     Make factor


Data modifications {#modifs}
-------------------

* Make `Survived` , `Pclass` and `Embark`   factors, 
* Create `Famly` = `SibSp + Parch` 

```{r datamod}
sf <- factor(tdf$Survived, labels=c("No", "Yes"))
# verification
# head(tdf$Survived)
# head(sf2)

tdf$Survived <- sf

# makefactor class
tdf$Pclass <- factor(tdf$Pclass)

# combine SibSp + Parch
tdf$Famly <- tdf$SibSp + tdf$Parch

tdf$Embarked[which(tdf$Embarked == "")] <- NA
tdf$Embarked <-  factor(tdf$Embarked)
levels(tdf$Embarked)

# ajout de variables
tdf <- tdf %>%
        mutate(sex.class = paste(Sex, Pclass, sep ="."))

```




Data Analysis {#analysis}
==================

Passenger Identity {#demographic}
-------------------

### Sex and Age {#sexage}

```{r ident1, warning=FALSE, fig.cap="Gender distribution"}

# opts.labels("f13"), fails

tbsex <- table(tdf$Sex)
pander(tbsex, caption = tabcap("Gender distribution"))

ggplot(tdf, aes(Sex)) + geom_bar()
```

```{r ident2, warning=FALSE, fig.cap="Age and Gender"}
sumAge <- summary(tdf$Age)
```

Numerical summaries of Age : `r pander(sumAge)`

```{r ident3,  warning=FALSE, fig.cap="Age and Gender"}
nc <- nclass.Sturges(tdf$Age)
nc <- nclass.FD(tdf$Age[!is.na(tdf$Age)])
ggplot(tdf, aes(Age)) + geom_histogram(bins=nc)
ggplot(tdf, aes(Age)) + geom_histogram(aes(y=..density..),bins=nc) + 
        facet_grid(Sex ~ .)

ggplot(tdf, aes(x=Sex, y=Age)) + geom_boxplot() +
        scale_x_discrete(limits = rev(levels(tdf$Sex))) +
        coord_flip() + scale_y_continuous(breaks = seq(0,90, by=10))
```




Family members on board {#fmly}
---------------


```{r family,  fig.cap="Family members on board", warning=FALSE}

pander(cor(tdf[ , c("SibSp","Parch","Famly")]), caption = tabcap("Family members: correlation of the variables"))
```

```{r family2, fig.cap="Family members on board", warning=FALSE}
ggplot(tdf) + geom_bar(aes(SibSp))
ggplot(tdf) + geom_bar(aes(Parch))
ggplot(tdf) + geom_bar(aes(Famly))

```

Passenger class {#pclass}
-------------------

```{r ship,  fig.cap="Passenger Classes"}
pander(prop.table(table(tdf$Pclass)), caption = tabcap("Class"))

ggplot(tdf) + geom_bar(aes(Pclass, fill = Pclass, group=1, y = ..prop.. ))
```

```{r family-class,  fig.cap="Passenger demographics by class"}

ggplot(tdf, aes(Sex)) + 
        geom_bar(aes(y = ..prop.., fill = Pclass, group=Pclass)) + 
        facet_grid(Pclass ~ .) +
        labs(y = "Proportion")

nc <- nclass.FD(tdf$Age[!is.na(tdf$Age)])
ggplot(tdf, aes(Age)) + 
        geom_histogram(aes(y=..density..,fill=Pclass),bins=nc) +
        facet_grid(Pclass ~ .)

ggplot(tdf, aes(Age)) + 
        geom_histogram(aes(y=..density..,fill=Pclass),bins=nc) +
        facet_grid(Pclass ~ Sex)


# ggplot(tdf) + geom_bar(aes(SibSp, y = ..prop..)) + facet_grid(Pclass ~ .)
# ggplot(tdf) + geom_bar(aes(Parch)) + facet_grid(Pclass ~ .)

ggplot(tdf) + 
        geom_bar(aes(Famly, y = ..prop.., fill=Pclass)) +
        facet_grid(Pclass ~ .)+ 
        scale_x_continuous(breaks=seq(0,10,by=2))

ggplot(tdf) + 
        geom_bar(aes(Famly, y = ..prop.., fill=Pclass, group=sex.class)) +
        facet_grid(Pclass ~ Sex) + 
        scale_x_continuous(breaks=seq(0,10,by=2))
```

Compare with reference: "r .ref(fig:ship)"


```{r Embarked}
table(tdf$Embarked)

sum(is.na(tdf$Embarked))

tdfx <- tdf[complete.cases(tdf), ]

ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., group=1))

ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., group=Sex)) + facet_grid(.~Sex)
ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., group=Pclass)) + facet_grid(Pclass~.)
ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., group=Pclass)) +
        facet_grid(Pclass~Sex)


```



```{r fare, w=w.12,a=a.11, fig.cap="Fare by class and Sex"}

# ncf <- nclass.FD(tdf$Fare) 
ncf <- 50
ggplot(tdf) + geom_histogram(aes(Fare), bins = ncf)
# ggplot(tdf) + geom_histogram(aes(Fare, y=..density.., fill=Pclass), bins = ncf) + 
#         facet_grid(Pclass ~.) 

ggplot(tdf) + geom_boxplot(aes(Pclass,Fare, fill = Pclass), color = "grey50") + 
        scale_x_discrete(limits = rev(levels(tdf$Pclass))) +
        coord_flip() +
        facet_grid(Sex ~.) 
```





Survival
--------

## overall

```{r survival, fig.cap="Overall survival"}
pander(prop.table(table(tdf$Survived)), digits=2)

ggplot(tdf) + geom_bar(aes(Survived, y = ..prop.., group = 1))
```

## By category

```{r survival-by}
# by sex
pander(prop.table(table(tdf$Survived, tdf$Sex),
                  margin = 2),
       digits=2)

# by class
pander(prop.table(table(tdf$Survived, tdf$Pclass),
                  margin = 2),
       digits=2)

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., group = Sex)) + 
        facet_grid(Sex ~.)

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., group = Pclass)) + 
        facet_grid(Pclass ~.) 


```

Class and sex

```{r survival-by2, fig.cap="Survival by Sex and Class"}

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Sex)
```


Class , sex, embarked

```{r survival-by3, fig.cap="Survival by Sex and Class / Embarked"}

ggplot(tdfx[tdfx$Embarked == "C", ]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Sex)+
        labs(title = "Embarked at Cherbourg")


ggplot(tdfx[tdfx$Embarked == "Q", ]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Sex)+
        labs(title = "Embarked at Queenstown")


ggplot(tdfx[tdfx$Embarked == "S", ]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Sex)+
        labs(title = "Embarked at Southampton")


```

Class and Age

```{r survival-by4, fig.cap="Age by Survival  and Class"}

ggplot(tdf) + 
        geom_histogram(aes(Age, y= ..density.., fill = Pclass), bins = nc) +
        facet_grid(Pclass ~ Sex) +
        labs(title= "Survivors")

ggplot(tdf[tdf$Survived == "Yes", ]) + 
        geom_histogram(aes(Age, y= ..density.., fill = Pclass), bins = nc) +
        facet_grid(Pclass ~ Sex) +
        labs(title= "Survivors")

ggplot(tdf[tdf$Survived == "No", ]) + 
        geom_histogram(aes(Age, y= ..density.., fill = Pclass), bins = nc) +
        facet_grid(Pclass ~ Sex) +
        labs(title= "Non-Survivors")
```






Two more lines, for just a try: 
Citing references: 

1. a figure: figure \@ref(fig:ship)
2. and tables: data head table (table \@ref(tab:dataget)) and summary table (table \@ref(tab:datasum)) and table \@ref(tab:mantable1))



** Bye-Bye **



