---
title: 'Titanic - Kaggle - Some exploration'
subtitle: "Explore for classroom use"
author:  "Bruno Fischer Colonimos"
date: "`r format(Sys.Date(), '%d %B %Y')`"
# header-includes: \usepackage{graphicx} # for unmodded template
header-includes: \usepackage{float} # for fig.pos = 'H'
fontsize: 12pt
urlcolor: "blue"   ## external
linkcolor: "red" ## internal
documentclass: article
classoption: a4paper
geometry: "left=1.5cm,right=1.5cm,top=1.5cm,bottom=2cm,footskip=1cm"
footer: \thetitle \ \textemdash \ \thepage 
output:
  # bookdown::pdf_document2:
  pdf_document:
    # template: test_latex_temp
    template: latex_temp2.latex
    highlight: monochrome
    number_sections: yes
    toc: yes
    toc_depth: 4
    # keep_tex: true
    fig_caption: yes
    keep_tex: yes
    
  # html_document:
  #   number_sections: yes
  #   toc: yes
  #   toc_depth: 4
  #   fig_caption: yes
---

***************************


<!-- Some Inline HTML (CSS) , formatting tables -->

<style>

table {
    width: 100%;
    border-top-style: solid;
    border-top-width: medium;
    border-top-color: initial;
    border-bottom-style: solid;
    border-bottom-width: medium;
    border-bottom-color: initial;
}

</style>

<!-- End Inline HTML -->



Preliminary code
================

```{r setup0, include=FALSE}

# Later
# =======
# knitr::opts_chunk$set(echo = TRUE, fig.keep="all" ,fig.width=4, fig.asp= 0.75 , fig.align="center", fig.keep ="all", fig.show="hold", out.width="0.48\\textwidth")

# , fig.align="center"
# not for HTML : out.width="0.48\\textwidth"
```


## Required packages (install before running)

"caret", "ggplot2", pander       



## Libraries and auxiliary code (install before running)

(Not fully echoed here)

```{r, echo = TRUE}
# adjust Gggplot2 theme and color palette
bwtheme <- TRUE # use ggplot2 bw -theme
specialpalette <- TRUE # controls the colour of the plots
showarning <- TRUE # show warnings (deleted empty rows...)
datmis <- "Ukn"
```



```{r libs_prep, echo =FALSE, warning=FALSE, message=FALSE, results='hide'}

library(knitr)
library(pander)
library(caret)
library(ggplot2)
library(dplyr)


# Source preparatory code
# ========================

source("preparation.R")

# Set aliases
knitr::set_alias(w = "fig.width", a = "fig.asp")

```

```{r setup}
# default options
knitr::opts_chunk$set(echo = FALSE, fig.width = w.13, fig.asp = a.11 , fig.pos = 'H', fig.align = "center", fig.show = "hold", warning = showarning)
```

Data
=====

Get it
-------

```{r dataget}
# read
datadir <-  "data"
fname <-  "train.csv"
tdf <- read.csv(file.path(datadir, fname), stringsAsFactors = FALSE)

```
 
 * The data. Table `r .ref("tab:datafirst")`

```{r datafirst}
pander(head(tdf), caption = tabcap("A glimpse of the data"))     
```

* Quick dataframe summary: table `r .ref("tab:datasum")`
```{r datasum}
pander(summary(tdf), caption = tabcap("Data summaries"))
```


Variables organization {#variables}
-----------------------


Variables                 type        Values                Treatment
------------------------  --------    -------------------   ------------------
__Demographic variables__
    * Sex                  String       female ; male       Make factor
    * Age                  numeric
    * Agestat              factor      enfant/ado/adulte    new:
__Family context__
    * SibSp                numeric                          Combined ->
    * Parch                numeric                          Combined ->
    * Famly                numeric                          new: SibSp + Parch
__Relationship to ship__
    * Pclass                             1,2,3              Make factor
    * ticket                            ticket number       not used
    * Fare                 numeric
    * Cabin                             Cabin nbr           modified -> Deck
    * Deck                 String                           new (from Cabin)
    * Embark               Factor       C = Cherbourg,          
                                        Q = Queenstown, 
                                        S = Southampton
__Survival__              
    * survived             binary        0/1                Make factor

:`r tabcap(chunklabel = "dataorg", "Data organization")`


Data modifications {#modifs}
-------------------

* Make `Survived` , `Pclass` and `Embark`   factors, 
* Create `Famly` = `SibSp + Parch` 


```{r datamod, warning=FALSE}
sf <- factor(tdf$Survived, labels=c("No", "Yes"))
tdf$Survived <- sf

# makefactor class
tdf$Pclass <- factor(tdf$Pclass)

# combine SibSp + Parch
tdf$Famly <- tdf$SibSp + tdf$Parch

# Embarked
tdf$Embarked[which(tdf$Embarked == "")] <- datmis
tdf$Embarked <-  factor(tdf$Embarked)
levels(tdf$Embarked)
# levels(tdf$Embarked)

# ajout de variables
tdf <- tdf %>%
        mutate(sex.class = paste(Sex, Pclass, sep ="."))


## Problemes résolus

tdf$agestat <- cut(tdf$Age, breaks=c(0,12,18,100), labels=c("child", "teen", "adult"))

tdf$agestat <- as.character(tdf$agestat)
tdf$agestat <- ifelse(is.na(tdf$agestat), datmis, tdf$agestat)
# tdf$agestat <- factor(tdf$agestat, levels = c("child", "teen", "adult", datmis))
tdf$agestat <- factor(tdf$agestat)
# levels(tdf$agestat)
# table(tdf$agestat)

# Ticket number begins with letters
tdf$letticket <- as.numeric(substr(tdf$Ticket,start = 1, stop = 1 ) %in% LETTERS)


# hascabin is the cabin number nkown?
tdf$hascabin <- as.numeric(tdf$Cabin != "")


# Deck (from https://triangleinequality.wordpress.com/2013/09/08/basic-feature-engineering-with-the-titanic-data/)

#This is going be very similar, we have a ‘Cabin’ column not doing much, only
#1st class passengers have cabins, the rest are ‘Unknown’. A cabin number looks
#like ‘C123’. The letter refers to the deck, and so we’re going to extract these
#just like the titles. Turning cabin number into Deck
#
# cabin_list <- c('A', 'B', 'C', 'D', 'E', 'F', 'T', 'G', 'Unknown')
# df[,'Deck']=df[,'Cabin'].map(lambda x: substrings_in_string(x, cabin_list))

tdf$Deck <- factor(ifelse( tdf$Cabin != "", 
                           substr(tdf$Cabin,start = 1,stop = 1), 
                           datmis))
tdf$Deck <- relevel(tdf$Deck, ref = "T")
# levels(tdf$Deck)

numna <- sapply(tdf,
       FUN = function(x){sum(is.na(x))} )

pander(t(numna), caption = "Number of missing values")
```




Data Analysis {#analysis}
==================

Passenger Identity {#demographic}
-------------------

### Sex and Age {#sexage}

```{r ident1, warning=FALSE, fig.cap="Gender distribution"}

# opts.labels("f13"), fails

tbsex <- table(tdf$Sex)
pander(tbsex, caption = tabcap("Gender distribution"))

ggplot(tdf, aes(Sex)) + geom_bar()
```

FIgure `r .ref("fig:ident1")`

```{r ident2, warning=FALSE, fig.cap="Age and Gender"}
sumAge <- summary(tdf$Age)
```

Numerical summaries of Age : `r pander(sumAge, caption = "Age distribution")`

```{r ident3,  warning=FALSE, fig.cap="Age and Gender"}
# nc <- nclass.Sturges(tdf$Age)
nc <- nclass.FD(tdf$Age[!is.na(tdf$Age)])
ggplot(tdf, aes(Age)) + geom_histogram(bins=nc)
ggplot(tdf, aes(Age)) + geom_histogram(aes(y=..density..),bins=nc) + 
        facet_grid(Sex ~ .)

ggplot(tdf, aes(x=Sex, y=Age)) + geom_boxplot() +
        scale_x_discrete(limits = rev(levels(tdf$Sex))) +
        coord_flip() + scale_y_continuous(breaks = seq(0,90, by=10))
```

FIgure `r .ref("fig:", "ident3")`


Family members on board {#fmly}
---------------


```{r family,  fig.cap="Family members on board", warning=FALSE}

pander(cor(tdf[ , c("SibSp","Parch","Famly")]), caption = tabcap("Family members: correlation of the variables"))
```

```{r family2, fig.cap="Family members on board", warning=FALSE}
ggplot(tdf) + geom_bar(aes(SibSp))
ggplot(tdf) + geom_bar(aes(Parch))
ggplot(tdf) + geom_bar(aes(Famly))
```

FIgure `r .ref("fig:", "family2")`

Passenger class {#pclass}
-------------------

figure `r .ref("fig:", "ship")`

```{r ship,  fig.cap="Passenger Classes Distribution"}

pander(local({
        tb <- rbind(table(tdf$Pclass),
                    prop.table(table(tdf$Pclass)))
        
        row.names(tb) = c("frequency", "rel.frequency")
        tb}), 
       caption = tabcap("Passenger Class (Pclass)"))



ggplot(tdf) + geom_bar(aes(Pclass, fill = Pclass, group=1, y = ..prop.. ))
```

```{r family_class, fig.cap="Passenger demographics by class"}

ggplot(tdf, aes(Sex)) + 
        geom_bar(aes(y = ..prop.., fill = Pclass, group=Pclass)) + 
        facet_grid(Pclass ~ .) +
        labs(y = "Proportion") +
        theme(legend.position = "none")

nc <- nclass.FD(tdf$Age[!is.na(tdf$Age)])
ggplot(tdf, aes(Age)) + 
        geom_histogram(aes(y=..density..,fill=Pclass),bins=nc) +
        facet_grid(Pclass ~ .) +
        theme(legend.position = "none")


ggplot(tdf, aes(Age)) + 
        geom_histogram(aes(y=..density..,fill=Pclass),bins=nc) +
        facet_grid(Pclass ~ Sex) +
        theme(legend.position = "none")



# ggplot(tdf) + geom_bar(aes(SibSp, y = ..prop..)) + facet_grid(Pclass ~ .)
# ggplot(tdf) + geom_bar(aes(Parch)) + facet_grid(Pclass ~ .)

ggplot(tdf) + 
        geom_bar(aes(Famly, y = ..prop.., fill=Pclass)) +
        facet_grid(Pclass ~ .)+ 
        scale_x_continuous(breaks=seq(0,10,by=2)) +
        theme(legend.position = "none")


ggplot(tdf) + 
        geom_bar(aes(Famly, y = ..prop.., fill=Pclass, group=sex.class)) +
        facet_grid(Pclass ~ Sex) + 
        scale_x_continuous(breaks=seq(0,10,by=2)) +
        theme(legend.position = "none")

```

FIgure `r .ref("fig:", "family_class")`
Compare with reference: `r .ref("fig:ship")`


```{r Deck}
ggplot(tdf) +
        geom_bar(aes(Deck, y=..prop.., group = 1))

ggplot(tdf[tdf$Deck != datmis,] ) +
        geom_bar(aes(Deck, y=..prop.., group = 1))

ggplot(tdf[tdf$Deck != datmis,] ) +
        geom_bar(aes(Deck, y=..prop.., fill = Pclass, group = Pclass)) + 
        facet_grid(Pclass ~ .) + theme(legend.position = "none")        
 
```


```{r Embarked, a= a.43, fig.cap="Embarkation port (C = Cherbourg, Q = Queenstown, S = Southampton)"}

temb <- table(tdf$Embarked)
ptemb <- prop.table(temb)
df <- rbind(frequency = temb , Rfreq = ptemb) 
pander(df, digits=2, caption = "Embarkation port")

na_emb <- sum(is.na(tdf$Embarked)) #  2 passengers have unknown embarkation port

tdfx <- tdf[complete.cases(tdf), ] 

ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., group=1))

ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., group=Sex)) + facet_grid(.~Sex) +
        theme(legend.position = "none")

ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., fill = Pclass,  group=Pclass)) + 
        facet_grid(Pclass~.) +
        theme(legend.position = "none")

ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., fill = Pclass, group=Pclass)) +
        facet_grid(Pclass~Sex) +
        theme(legend.position = "none")

ggplot(tdfx) + geom_bar(aes(Embarked, y=..prop.., fill = Pclass, group=Deck)) +
        facet_grid(Deck~Sex) +
        theme(legend.position = "none")
```



```{r fare, w=w.12,a=a.11, fig.cap="Fare by class and Sex"}

# ncf <- nclass.FD(tdf$Fare) 
ncf <- 50
ggplot(tdf) + geom_histogram(aes(Fare), bins = ncf)
# ggplot(tdf) + geom_histogram(aes(Fare, y=..density.., fill=Pclass), bins = ncf) + 
#         facet_grid(Pclass ~.) 

ggplot(tdf) + geom_boxplot(aes(Pclass,Fare, fill = Pclass), color = "grey50") + 
        scale_x_discrete(limits = rev(levels(tdf$Pclass))) +
        coord_flip() +
        facet_grid(Sex ~.) 
```


Survival
--------

## overall

```{r survival, fig.cap="Overall survival"}
pander(prop.table(table(tdf$Survived)), digits=2)

ggplot(tdf) + geom_bar(aes(Survived, y = ..prop.., group = 1))
```

## By category

```{r survival-by, fig.cap="Survival by sex or Passenger Class"}
# by sex
pander(prop.table(table(tdf$Survived, tdf$Sex),
                  margin = 2),
       digits=2)

# by class
pander(prop.table(table(tdf$Survived, tdf$Pclass),
                  margin = 2),
       digits=2)

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., group = Sex)) + 
        facet_grid(Sex ~.)

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., group = Pclass)) + 
        facet_grid(Pclass ~.) 


```

Class and sex

```{r survival-by2, fig.cap="Survival by Sex and Class"}

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Sex) +
        theme(legend.position = "none")
```


Class , sex, embarked

```{r survival-by3, fig.cap="Survival by Sex and Class / Embarked"}

ggplot(tdfx[tdfx$Embarked == "C", ]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Sex)+
        labs(title = "Embarked at Cherbourg")+
        theme(legend.position = "none")


ggplot(tdfx[tdfx$Embarked == "Q", ]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Sex)+
        labs(title = "Embarked at Queenstown")+
        theme(legend.position = "none")


ggplot(tdfx[tdfx$Embarked == "S", ]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ Sex)+
        labs(title = "Embarked at Southampton")+
        theme(legend.position = "none")


```

Class and Age see figure `r .ref("fig:survival-by4")`

```{r survival-by4, fig.cap="Age by Survival  and Class", warning=FALSE}

ggplot(tdf) + 
        geom_histogram(aes(Age, y= ..density.., fill = Pclass), bins = nc) +
        facet_grid(Pclass ~ Sex) +
        labs(title= "All") + 
        theme(legend.position = "none")
        

ggplot(tdf[tdf$Survived == "Yes", ]) + 
        geom_histogram(aes(Age, y= ..density.., fill = Pclass), bins = nc) +
        facet_grid(Pclass ~ Sex) +
        labs(title= "Survivors") + 
        theme(legend.position = "none")

ggplot(tdf[tdf$Survived == "No", ]) + 
        geom_histogram(aes(Age, y= ..density.., fill = Pclass), bins = nc) +
        facet_grid(Pclass ~ Sex) +
        labs(title= "Non-Survivors") + 
        theme(legend.position = "none")
```

Class and agestat, agestat and sex

```{r survival-by5, fig.cap="Survival by age status, Class and sex"}

agesex_class <- table(tdf$agestat, tdf$Sex, tdf$Pclass)
pander(agesex_class, caption = "Distribution of age-status vs sex and Pclass" )




ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ agestat) + 
        labs(title = "Sex = All") +
        theme(legend.position = "none")

ggplot(tdf[tdf$Sex == "female" ,]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ agestat) +
        labs(title = "Sex ='female'") + 
        theme(legend.position = "none")

ggplot(tdf[tdf$Sex == "male" ,]) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Pclass, fill = Pclass))+ 
        facet_grid(Pclass ~ agestat)+
        labs(title = "Sex = 'male'") + 
        theme(legend.position = "none")

```


```{r survival_by6, w = w.23, fig.asp = 2, fig.cap="Survival by Deck, age and Sex"}

ggplot(tdf) + 
        geom_bar(aes(Survived, y = ..prop.., 
                     group = Deck, fill = Pclass))+ 
        facet_grid(Deck ~ agestat)+
        labs(title = "by Deck") # + theme(legend.position = "none")
```



